---
title: "Food recommender interface"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: fill
    source_code: embed
runtime: shiny
---

```{r setup, include=FALSE}
library(Matrix)
library(shiny)
library(flexdashboard)
library(DT)
library(rmarkdown)
library(stopwords)
library(SnowballC)
library(purrr)
library(tidyverse)

library(OutlierIdentifiers)
library(SparseMatrixRecommender)
library(SMRMon)
library(LSAMon)
```

```{r get-data, include=FALSE}

  load( file = "smrFineliFoods.RData" )
  load( file = "lsaFineliFoodWordsAndPhrases.RData" )
  load( file = "dfFineliFoodData.RData" )
  
  load( file = "smrUSDAFoods.RData" )
  load( file = "lsaUSDAFoodWordsAndPhrases.RData" )
  load( file = "dfUSDAFoodData.RData" )
  
  lsRecommenders <- list( "Fineli" = smrFineliFoods,               "USDA" = smrUSDAFoods )
  lsLSAObjects   <- list( "Fineli" = lsaFineliFoodWordsAndPhrases, "USDA" = lsaUSDAFoodWordsAndPhrases )
  lsData         <- list( "Fineli" = dfFineliFoodData,             "USDA" = dfUSDAFoodData )
  
  lsRecommenders <- 
    purrr::map( lsRecommenders, function(smr) {
      smr %>% SMRMonApplyTermWeightFunctions( "IDF", "None", "Cosine" )
    })
```

Parameters {.sidebar}
=======================================================================

```{r}
selectInput( inputId = "recommender", 
             label = "Type:", 
             choices = setNames( names(lsRecommenders), paste( names(lsRecommenders), "recommender" ) ), 
             selected = "Fineli")

sliderInput( inputId = "numberOfRecs", label = "Number of recommendations:", min = 1, max = 300, value = 10, step = 1 )

#sliderInput( inputId = "titleTagTypeWeight", label = "Title weight:", min = 0, max = 1, value = 1, step = 0.01 )

sliderInput( inputId = "wordTagTypeWeight", label = "Word weight:", min = 0, max = 1, value = 1, step = 0.01 )

sliderInput( inputId = "topicTagTypeWeight", label = "Topic weight:", min = 0, max = 1, value = 1, step = 0.01 )

```

```{r}
smrObj <- 
  reactive({
    lsRecommenders[[ input$recommender ]]
  })

lsaObj <- 
  reactive({
    lsLSAObjects[[ input$recommender ]]
  })

dfData <- 
  reactive({
    lsData[[ input$recommender ]]
  })
```

Main
=======================================================================

Row {data-height=800}
-----------------------------------------------------------------------

### Tags

```{r}
textAreaInput( 
  inputId = "tagsText", 
  label = "Tag text (metadata)", 
  placeholder = "FUCLASS:MEATDISH, Word:chicken", 
  width = "600px", cols = 300, rows = 8 )

lsTagsProf <- 
  reactive({
    
    if( nchar(input$tagsText) == 0) { 
      NULL 
    } else {
      res <- trimws( strsplit(input$tagsText, split = ",")[[1]] )
      res <- res[ nchar(res) > 0 ]
      res <- res[ res %in% colnames(smrObj()$M) ] 
    }
  })
```

Known tags:

```{r}
renderText(paste(lsTagsProf(), collapse = ", "))
```

### Free text

```{r}
textAreaInput( 
  inputId = "freeText", 
  label = "Free text (description)", 
  placeholder = "Hamburger with mayonnaise and tomatoes.", 
  width = "600px", cols = 300, rows = 8 )

lsFreeWordsProf <- 
  reactive({
    
    if( nchar(input$freeText) == 0) { 
      NULL 
    } else {
      
      matRes <- 
        lsaObj() %>% 
        LSAMonRepresentByTerms( query = input$freeText ) %>% 
        LSAMonTakeValue
      
      lsCSums <- colSums(matRes)
      lsProf <- lsCSums[lsCSums>0]
      
      names(lsProf) <- paste0( "Word:", names(lsProf) )
      lsProf <- lsProf[ names(lsProf) %in% colnames(smrObj()$M) ]
      rev(sort(lsProf))
    }
    
  })

lsFreeTopicsProf <- 
  reactive({
    
    if( nchar(input$freeText) == 0) { 
      NULL 
    } else {
      
      matRes <- 
        lsaObj() %>% 
        LSAMonRepresentByTopics( query = input$freeText ) %>% 
        LSAMonTakeValue
      
      matRes <- SMRApplyTermWeightFunctions( matRes, "None", "None", "Cosine" )
      
      lsCSums <- colSums(matRes)
      lsProf <- lsCSums[lsCSums>0]
      
      names(lsProf) <- paste0( "Topic:", names(lsProf) )
      lsProf <- lsProf[ names(lsProf) %in% colnames(smrObj()$M) ]
      rev(sort(lsProf))
    }
    
  })
```


Extracted words and topics:

```{r}
lsFreeProf <- reactive({ c( lsFreeWordsProf(), lsFreeTopicsProf() ) })
DT::renderDataTable(
  expr = data.frame( Word = names(lsFreeProf()), Score =  lsFreeProf() ),
  options = list(scrollX = "300px", scrollY = "300px", scroller = TRUE)
)
```

Row
-----------------------------------------------------------------------


### Recommendations with both tags and free text

```{r}
lsProf <- 
  reactive({ 
    c( 
      setNames( rep_len( x = 1, length.out = length(lsTagsProf())), lsTagsProf() ), 
      lsFreeProf() 
    ) 
  })

DT::renderDataTable(
  if( is.null(lsProf()) || length(lsProf()) == 0 ) {
    NULL
  } else {
    smrObj() %>% 
      SMRMonApplyTagTypeWeights( 
        weights = 
          c( 
            Word = input$wordTagTypeWeight, 
            Topic = input$topicTagTypeWeight
          )[ smrObj() %>% SMRMonTakeTagTypes ], 
        default = 1 ) %>% 
      SMRMonRecommendByProfile( profile = lsProf(), nrecs = input$numberOfRecs ) %>% 
      SMRMonJoinAcross( dfData() ) %>% 
      SMRMonTakeValue
    },
  options = list(scrollX = "300px", scrollY = "300px", scroller = TRUE)
) 
```

Summary
=======================================================================

### Dimensions of the recommender matrix

```{r}
renderText( dim(smrObj() %>% SMRMonTakeM) )
```

### Tag type ranges

```{r}
DT::renderDataTable( smrObj() %>% SMRMonTakeTagTypeRanges, 
                     options = list( pageLength = 20 ) )
```

### Tags statistics

Row
----------------

```{r}
# uiOutput("summaryControl")
# 
# output$summaryControl <- 
#   renderUI({
#     selectInput( inputId = "summaryTagType", 
#                  label = "Tag type:", 
#                  choices = smrObj() %>% SMRMonTakeTagTypes, 
#                  selected =  (smrObj() %>% SMRMonTakeTagTypes)[[1]] )
#     
#   })

dfSummaryTagTypeTags <- 
  reactive({
    res <- colSums( SMRUnitize( smrObj() %>% SMRMonTakeM ) )
    res <- data.frame( Tag = names(res), Frequency = setNames(res, NULL) )
    res <- cbind( res, TagType = sapply( strsplit(res$Tag, ":"), function(x) x[[1]][[1]] ) )
    res %>% 
      dplyr::arrange( TagType, desc(Frequency) ) %>% 
      dplyr::select( TagType, Tag, Frequency)
  })

DT::renderDataTable( 
  dfSummaryTagTypeTags(),
  options = list(pageLength = 20)
) 
```


References
=======================================================================

#### In brief

The data of the Fineli recommender was taken from
[Fineli Open Data](https://fineli.fi/fineli/en/avoin-data), [FI1].

That data was put in the R data package [FineliFoodData](https://github.com/antononcube/R-packages/tree/master/FineliFoodData)

The data for the USDA recommender was taken from
[NutrienTrackeR](https://CRAN.R-project.org/package=NutrienTrackeR), [ARp1].

The recommenders were made with the packages 
[`SMRMon-R`](https://github.com/antononcube/R-packages/tree/master/SMRMon-R)
and
[`LSAMon-R`](https://github.com/antononcube/R-packages/tree/master/LSAMon-R),
[AAp1, AAp2, AAp3].


#### Data

[FI1] National Institute for Health and Welfare, Fineli,
[Fineli food database](https://fineli.fi/fineli/en/avoin-data),
https://fineli.fi.


#### Packages 

[ARp1] Andrea Rodriguez-Martinez, Rafael Ayala, Mark Balchunas, Yacine Debbabi, Lara Selles Vidal,
[NutrienTrackeR](https://CRAN.R-project.org/package=NutrienTrackeR),
(2021),
[CRAN](https://github.com/AndreaRMICL).

[ARr1] Andrea Rodriguez-Martinez, Rafael Ayala, Mark Balchunas, Yacine Debbabi, Lara Selles Vidal,
[NutrienTrackeR](https://github.com/AndreaRMICL/NutrienTrackeR),
(2018),
[GitHub/AndreaRMICL](https://github.com/AndreaRMICL).

[AAp1] Anton Antonov,
[Sparse Matrix Recommender framework functions, R-package](https://github.com/antononcube/R-packages/tree/master/SparseMatrixRecommender),
(2019),
[R-packages at GitHub/antononcube](https://github.com/antononcube/R-packages).

[AAp2] Anton Antonov,
[Sparse Matrix Recommender Monad, R-package](https://github.com/antononcube/R-packages/tree/master/SMRMon-R),
(2019),
[R-packages at GitHub/antononcube](https://github.com/antononcube/R-packages).

[AAp3] Anton Antonov,
[Latent Semantic Analysis Monad, R-package](https://github.com/antononcube/R-packages/tree/master/LSAMon-R),
(2019),
[R-packages at GitHub/antononcube](https://github.com/antononcube/R-packages).


