---
title: "Fineli Food Data Recommenders"
author: Anton Antonov
date: 2021-03-28
output: html_notebook
params:
  saveDirName: "../output/"
  exportDirName: "../smr-export/"
  saveObjectsQ: TRUE
  exportSMRQ: TRUE
---

```{r}
library(d3heatmap)
library(magrittr)
library(tidyverse)
library(Matrix)
library(SparseMatrixRecommender)
library(SMRMon)
library(LSAMon)
library(OutlierIdentifiers)
library(SparseMatrixRecommenderInterfacesNoDT)
library(ParetoPrincipleAdherence)

library(FineliFoodData)
```


# Introduction

There are three recommenders that can be made with the Fineli Food Data:

- [X] Overall recommender

- [ ] Recipes recommender

- [ ] "Simple" foods recommender

We first make the "Overall recommender" and then enhance it with foodID-to-foodID to matrix that contains 
the relationships between food recipes and food ingredients.

# Data preparation

```{r}
dfFineliFoodDataOrig <- lsFineliFoodDataTables$food
head(dfFineliFoodDataOrig)
```


# Overall recommender

```{r}
smrFineliFoodsFreq <- 
  SMRCreate( 
    dataRows = lsFineliFoodDataTables$food, 
    tagTypes = names(lsFineliFoodDataTables$food)[-c(1,2)],
    itemColumnName = "FOODID", 
    addTagTypesToColumnNamesQ = TRUE )
smrFineliFoodsFreq %>% SMRMonTakeTagTypeRanges
```


```{r}
smrFineliFoods <- 
  smrFineliFoodsFreq %>% 
  SMRMonApplyTermWeightFunctions( "IDF", "None", "Cosine" )
```

# Food names (single-)words topics

```{r}
set.seed(665)
lsaFoodNames <- 
  LSAMonUnit( iconv(MakeDictionary( "foodname", lang = "EN")) ) %>% 
  LSAMonMakeDocumentTermMatrix(stemWordsQ = FALSE, stopWords = stopwords::stopwords() ) %>% 
  LSAMonApplyTermWeightFunctions( "IDF", "None", "Cosine" ) %>% 
  LSAMonExtractTopics( numberOfTopics = 40, minNumberOfDocumentsPerTerm = 3, method = "NNMF", maxSteps = 15)
```

```{r rows.print=20}
lsaFoodNames <- 
  lsaFoodNames %>% 
  LSAMonEchoTopicsTable(numberOfTerms = 20, wideFormQ = T)
```


# Words and phrases topics

```{r}
lsaFoodPhrases <- 
  LSAMonUnit( iconv(MakeDictionary( "foodname", lang = "EN")) ) %>% 
  LSAMonMakeDocumentTermMatrix(stemWordsQ = FALSE, stopWords = NULL, splitPattern = ",\\s*" )
```

```{r}
sample(colnames(lsaFoodPhrases %>% LSAMonTakeDocumentTermMatrix),100)
```

```{r}
smat1 <- lsaFoodNames %>% LSAMonTakeDocumentTermMatrix
smat2 <- lsaFoodPhrases %>% LSAMonTakeDocumentTermMatrix
smat2 <- smat2[, setdiff(colnames(smat2), colnames(smat1))]
lsaFineliFoodWordsAndPhrases <- 
  LSAMonUnit() %>% 
  LSAMonSetDocumentTermMatrix( documentTermMatrix = cbind( smat1, smat2 ) ) %>% 
  LSAMonApplyTermWeightFunctions( "IDF", "None", "Cosine" ) %>% 
  LSAMonExtractTopics( numberOfTopics = 40, minNumberOfDocumentsPerTerm = 3, method = "NNMF", maxSteps = 15)
```

```{r rows.print=20}
lsaFineliFoodWordsAndPhrases <- 
  lsaFineliFoodWordsAndPhrases %>% 
  LSAMonEchoTopicsTable(numberOfTerms = 20, wideFormQ = T)
```
```{r, rows.print = 20, message=FALSE}
dfWordsNNs <- 
  lsaFineliFoodWordsAndPhrases %>% 
  LSAMonExtractStatisticalThesaurus( 
    searchWords = ( c("alaska", "game", "chicken", "salmon", "yogurt", "milk", "coffee", "pizza chain", "french", "vegetable soup")), 
    numberOfNearestNeighbors = 12, 
    fixed = T) %>% 
  LSAMonTakeValue %>% 
  dplyr::bind_rows()
purrr::map_df( split(dfWordsNNs, dfWordsNNs$SearchTerm), function(x) data.frame( SearchTerm = x$SearchTerm[[1]], Thesaurus = paste(x$Word.Word, collapse = ", ") ) )
```


# Extend core food recommender

```{r}
matWords <- lsaFoodWordsAndPhrases %>% LSAMonTakeWeightedDocumentTermMatrix
matTopics <- lsaFoodWordsAndPhrases %>% LSAMonNormalizeMatrixProduct(normalizeLeftQ = F) %>% LSAMonTakeW
```

```{r}
smrFineliFoods <- 
  smrFineliFoods %>% 
  SMRAnnexSubMatrix( newSubMat = matWords, newTagType = "Word", imposeSameRowNamesQ = T, addTagTypesToColumnNamesQ = T) %>% 
  SMRAnnexSubMatrix( newSubMat = matTopics, newTagType = "Topic", imposeSameRowNamesQ = T, addTagTypesToColumnNamesQ = T) %>% 
  SMRMonApplyTermWeightFunctions( "None", "None", "Cosine")
```

# Make interpretable data

```{r}
smrFineliFoods %>% SMRMonTakeTagTypeRanges
```

```{r}
colnames(SMRSubMatrix(smrFineliFoods, "IGCLASS"))
```

```{r}
lsFineliFoodDataTables$foodname_EN
```


```{r}
aFoodName <- MakeDictionary("foodname", nameFrom = "FOODID", nameTo = "FOODNAME", lang = "EN")
aProcess <- MakeDictionary("process", nameFrom = "THSCODE", nameTo = "DESCRIPT", lang = "EN")
aIgClass <- MakeDictionary("igclass", nameFrom = "THSCODE", nameTo = "DESCRIPT", lang = "EN")
aFuClass <- MakeDictionary("fuclass", nameFrom = "THSCODE", nameTo = "DESCRIPT", lang = "EN")
```

```{r}
head(dfFineliFoodDataOrig)
```

```{r}
dfFineliFoodData <- 
  dfFineliFoodDataOrig %>% 
  dplyr::mutate( FOODID = as.character(FOODID) ) %>% 
  dplyr::mutate( FOODNAME = aFoodName[FOODID]  ) %>% 
  dplyr::mutate( PROCESS  = aProcess[PROCESS]  ) %>% 
  dplyr::mutate( FUCLASS  = aFuClass[FUCLASS]  ) %>% 
  dplyr::mutate( FUCLASSP = aFuClass[FUCLASSP] ) %>% 
  dplyr::mutate( IGCLASS  = aIgClass[IGCLASS]  ) %>% 
  dplyr::mutate( IGCLASSP = aIgClass[IGCLASSP] )
dfFineliFoodData
```

```{r}
Summary(dfFineliFoodData)
```

# Re-map recomemnder items

```{r}
aFoodIDToName <- MakeDictionary( "foodname", lang = "EN", columnNamePrefixesQ = F)
aFoodNameToID <- setNames( names(aFoodIDToName), aFoodIDToName )
```

```{r, eval=F}
rownames(smrFineliFoodsFreq$M) <- aFoodIDToName[rownames(smrFineliFoodsFreq$M) ]
rownames(smrFineliFoodsFreq$M01) <- aFoodIDToName[rownames(smrFineliFoodsFreq$M01) ]
```

```{r, eval=F}
rownames(smrFineliFoods$M) <- aFoodIDToName[rownames(smrFineliFoods$M) ]
rownames(smrFineliFoods$M01) <- aFoodIDToName[rownames(smrFineliFoods$M01) ]
```

# Example recommendations

```{r}
sample(aFoodIDToName, 3)
```




```{r, rows.print = 20}
smrFineliFoods %>% 
  SMRMonRecommend( history = aFoodNameToID[ "BLACKCURRANT JUICE, UNSWEETENED, DILUTED, HOME-MADE"], nrecs = 20) %>% 
  SMRMonJoinAcross( data = dfFineliFoodData, by = "FOODID") %>% 
  SMRMonTakeValue
```


```{r}
grep("Word:m.*ss", colnames(smrFineliFoods$M), value = T)
```

```{r, rows.print = 20}
query <- "moussaka with potatoes"
matProf <- lsaFoodNames %>% LSAMonRepresentByTerms( query = query ) %>% LSAMonTakeValue
lsProf <- colnames(matProf)[ colSums(matProf) > 0];
lsProf <- paste0( "Word:", lsProf)
lsProf
```

```{r, rows.print = 20}
smrFineliFoods %>% 
  SMRMonRecommendByProfile( profile = lsProf, nrecs = 20) %>% 
  SMRMonJoinAcross( dfFineliFoodData ) %>% 
  SMRMonTakeValue
```

```{r}
Summary(dfFineliFoodDataOrig, maxsum = 20)
```

```{r, rows.print = 20}
smrFineliFoods %>% 
  SMRMonApplyTagTypeWeights( weights = c("Word" = 0, "Topic" = 0), default = 1) %>% 
  SMRMonRecommend( history = aFoodNameToID["MOUSSAKA, PORK-BEEF MINCE, POTATOES, LOW-FAT MILK"], nrecs = 20) %>% 
  SMRMonJoinAcross( dfFineliFoodDataOrig ) %>% 
  SMRMonTakeValue
```

# Save

```{r}
if( params$saveObjectsQ ) {
  save( smrFineliFoods,               file = file.path( params$saveDirName, "smrFineliFoods.RData"))
  save( lsaFineliFoodWordsAndPhrases, file = file.path( params$saveDirName, "lsaFineliFoodWordsAndPhrases.RData"))
  save( dfFineliFoodData,             file = file.path( params$saveDirName, "dfFineliFoodData.RData"))
}
```

# Export

```{r}
if( params$exportSMRQ ) {
  SMRExportToDirectory( smr = smrFineliFoods,  directoryPath = params$exportDirName, dataNameInfix = "FineliFoods" )
}
```
