---
title: "Fineli data study"
author: Anton Antonov
date: 2021-03-27
output: html_notebook
params:
  fineliDataDirName: "~/Datasets/Fineli/Fineli_Rel20"
  fineliLanguage: "EN"
---

```{r setup}
library(d3heatmap)
library(magrittr)
library(tidyverse)
library(Matrix)
library(SparseMatrixRecommender)
library(SMRMon)
library(LSAMon)
library(OutlierIdentifiers)
library(SparseMatrixRecommenderInterfacesNoDT)
library(ParetoPrincipleAdherence)

library(FineliFoodData)
```

# Introduction

In this notebook we study the transitive relationships between the "Finnish Food Dataset" [Fineli]( https://fineli.fi).

The Fineli data has a fairly complete description of its structure and organization. See the documents:

- "Fineli®, food composition data files"

- "Nutrient factors included in Fineli packages"

The data is three different languages: English, Finnish, Swedish.

**Remark:** This notebook is parameterized with the focus language. Here is the correspondence between language and parameter value:
 English (EN), Finnish (FI), Swedish (SV).

# Analyze relationships

```{r}
lsTableColnames <- purrr::map( lsFineliFoodDataTables, function(x) colnames(x) )
dfPairs <- expand.grid( names(lsTableColnames), names(lsTableColnames) )
dfPairs <- dfPairs[dfPairs[[1]] != dfPairs[[2]], ]
dim(dfPairs)
```


```{r}
head(dfPairs)
```


```{r}
dfCommonColnamesLongForm <- 
  purrr::map_df( split(dfPairs, 1:nrow(dfPairs)), function(dfX) {
    cbind( dfX,
           NCommon = length( intersect(lsTableColnames[[ dfX[1,1] ]], lsTableColnames[[ dfX[1,2] ]] ) ),
           Common = paste( intersect(lsTableColnames[[ dfX[1,1] ]], lsTableColnames[[ dfX[1,2] ]]), collapse = "; "  ) 
    )
  })
```


```{r}
dfCommonColnamesLongForm %>% dplyr::sample_n(10)
```

```{r}
xtabs( NCommon ~ Var1 + Var2, dfCommonColnamesLongForm, sparse = F ) 
```

```{r, rows.print=20}
dfCommonColnames <- dfCommonColnamesLongForm %>% pivot_wider( id_cols = "Var1", names_from = "Var2", values_from = "Common", values_fn = as.character)
View(dfCommonColnames)
```

# Connections investigations

## Connection between recipes and foods

Here is the summary of the table "food":

```{r}
Summary(lsFineliFoodDataTables$food)
```
We can see that food names have 1-to-1 correspondence with food IDs. (Hence can be used as items/rownames in the SMR object.)

Here is the summary of the table "contribfood":

```{r}
Summary(lsFineliFoodDataTables$contribfood)
```

The recipes table "contribfood" is in long form. The ID column is "FOODID", the Variable column is "CONFDID".

With this summary we see that most of recipes are "non-ingredient" foods ("IGCLASSP:NONINGR") :

```{r}
Summary(lsFineliFoodDataTables$food %>% 
          dplyr::mutate( FOODNAME = NULL ) %>% 
          dplyr::right_join( lsFineliFoodDataTables$contribfood %>% dplyr::select( FOODID ), by = c("FOODID"="FOODID") ) %>% 
          dplyr::inner_join( lsFineliFoodDataTables$foodname_EN, by = "FOODID" ) %>% 
          dplyr::mutate( LANG = NULL ) ) 
```
Let us compare with the summary of the corresponding anti-join:

```{r}
Summary(lsFineliFoodDataTables$food %>% 
          dplyr::mutate( FOODNAME = NULL ) %>% 
          dplyr::right_join( lsFineliFoodDataTables$contribfood %>% dplyr::select( FOODID ), by = c("FOODID"="FOODID") ) %>% 
          dplyr::inner_join( lsFineliFoodDataTables$foodname_EN, by = "FOODID" ) %>% 
          dplyr::mutate( LANG = NULL ) ) 
```



```{r}
lsFineliFoodDataTables$food %>% 
  dplyr::mutate( FOODNAME = NULL ) %>% 
  dplyr::inner_join( lsFineliFoodDataTables$foodname_EN, by = "FOODID" ) %>% 
  dplyr::inner_join( lsFineliFoodDataTables$fuclass_EN, by = c( "FUCLASS" = "THSCODE" ) ) %>% 
  dplyr::inner_join( lsFineliFoodDataTables$igclass_EN, by = c( "IGCLASS" = "THSCODE" ) ) %>% 
  dplyr::inner_join( lsFineliFoodDataTables$fuclass_EN, by = c( "FUCLASSP" = "THSCODE" ) ) %>% 
  dplyr::inner_join( lsFineliFoodDataTables$igclass_EN, by = c( "IGCLASSP" = "THSCODE" ) ) %>% 
  dplyr::select( !tidyselect::starts_with("LANG")) %>% 
  dplyr::select( !tidyselect::starts_with("DESC")) %>% 
  dplyr::inner_join( lsFineliFoodDataTables$process_EN, by = c( "PROCESS" = "THSCODE" ) )
```

```{r}
lsFineliFoodDataTables$contribfood %>% 
  dplyr::inner_join( lsFineliFoodDataTables$foodname_EN %>% dplyr::mutate( LANG = NULL, INGREDIENTFOODNAME = FOODNAME, FOODNAME = NULL ), 
                     by = c("CONFDID"="FOODID") ) %>% 
  dplyr::inner_join( lsFineliFoodDataTables$foodname_EN %>% dplyr::mutate( LANG = NULL ), 
                     by = c("FOODID"="FOODID") )
```
# Other studies

```{r}
lsFineliFoodDataTables$foodname_EN
```

```{r}
dfCountryAdjectiveToName <- read.csv("https://raw.githubusercontent.com/antononcube/Raku-DSL-Entity-English-Geographics/main/resources/CountryAdjectiveToName.csv", header = FALSE)
names(dfCountryAdjectiveToName) <- c("Adjective", "Name")
dim(dfCountryAdjectiveToName)
```

```{r}
lsHasCountrySpec <- 
  purrr::map_lgl( iconv(lsFineliFoodDataTables$foodname_EN$FOODNAME), function(x) {
    length(intersect( tolower(strsplit(x, " ")[[1]]), tolower(dfCountryAdjectiveToName$Adjective))) > 0 ||
          length(intersect( tolower(strsplit(x, " ")[[1]]), tolower(dfCountryAdjectiveToName$Name))) > 0
  })
mean(lsHasCountrySpec)
```


```{r, rows.print = 30}
cbind( lsFineliFoodDataTables$foodname_EN, HasCountrySpec = lsHasCountrySpec ) %>% 
  dplyr::filter( HasCountrySpec )
```

